---
title: "tibbleColumns"
author: "Hoyt Emerson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r message=FALSE,echo=FALSE}
library(tidyverse)
library(tibbleColumns)
```
The tibbleColumns package is designed to offer some time saving functions that fix problems you didn't even know you had within the tidyverse. It also introduce some advanced methods I've developed to use in your pipe sequences. Some examples of functions this package offers is:

- Quick proportion columns
- Quick x_o_x columns for MoM, YoY etc.
- Perform t-tests with tibbles instead of coverting to data.frame
- Get a nice lm summary tibble a la broom package
- "Tibble Out" the current state of a pipe sequence to your global environment
- Plug in a "Tibble Module" to manipulate data outside the pipe sequence

## Proportion Columns

This entire package started when I realized that creating a proportion column in R was not a simple and easy thing for tibbles. I was frustrated that Excel was able to do something better than R so I decided to do something about it.

```{r, fig.show='hold'}
#prop_column_group
mtcars %>% prop_column_group(cyl)

#prop_column
mtcars %>% count(cyl, mpg) %>% prop_column(n)
```

## XoX Columns

There are two particular functios in the tibbleColumns package that create either a column representing the percentage change from one number to another or an entirely new tibble with a group, number and percentage change.

The **change_XoX_column** creates a single column that calculates the percentage of change between the vectors in two other columns. These columns need to be organized from left to right. Meaning, column of prior month data to the left of a column for current month.

```{r, fig.show='hold'}
#change_XoX_column
tc <- tibble(
  Month = c("Jan", "Dec"),
  Users = c(102, 909)
)

tc %>% spread(Month, Users) %>% change_XoX_column(Dec, Jan, "MoM")
```

The **change_XoX_column_group** will simply look at an entire data set and do all the above coding for you along with adding a group column as well.

So let’s say you want to find out the MoM change in revenue by device from our hypothetical data set. The function will do that very quickly:

<em>NOTE: For this function you need to select certain columns of the data in a certain order. The columns also need to be specific data types. The order goes Group (character), Date (character), Aggregate Number (numeric)</em>

```{r, fig.show='hold'}
#change_XoX_column_group
tcg <- tibble(
  Month = c("Jan", "Jan","Dec", "Dec"),
  Type = c("Red", "Blue", "Red", "Blue"),
  Users = c(102, 909, 201, 479)
)

tcg %>% select(Type, Month, Users) %>% change_XoX_column_group(Dec, Jan, "MoM")
```

## T-test

The tibble interacts poorly with base stat functions like t.test(). Instead of just complaining about how data_frame() wasn’t recognized by the base function, I decided to create a wrapper function that simply does all the dirty work for you. So you can now pass two tibbles to this function and it will return a nice tibble summary of your t-test.

```{r, fig.show='hold'}
#testoutput
t1 <- tibble(
    type = c("Blue", "Blue", "Blue", "Blue"),
    num = runif(4,0,5)
  )

t2 <- tibble(
  type = c("Red","Red", "Red", "Red"),
  num = runif(4,0,5)
)

ttest_tibble(t1$num,t2$num)
```


## 

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
